An Improved Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium

### 2021-04-06 simplified_sdc_intro/paper.tex
add NSE bailout description

@@

@@

\begin{deluxetable}{lllllll}
\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting with $T$ evolution disabled.}
\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
\startdata
 $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
 $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
 $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
 $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
 $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
 $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
 $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
 $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
 $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
 $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
\enddata
\end{deluxetable}

% run on 12-31-20
% Castro: 20.12-80-gfe3e9c8de8
\begin{deluxetable}{lllllll}
\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
\startdata
 $\rho$                      & $2.794 \times 10^{18}$  & 2.047  & $6.760 \times 10^{17}$  & 2.563  & $1.144 \times 10^{17}$  \\
 $\rho u$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
 $\rho v$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
 $\rho E$                    & $2.450 \times 10^{35}$  & 2.357  & $4.783 \times 10^{34}$  & 2.743  & $7.144 \times 10^{33}$  \\
 $\rho e$                    & $2.261 \times 10^{35}$  & 2.325  & $4.513 \times 10^{34}$  & 2.821  & $6.385 \times 10^{33}$  \\
 $T$                         & $2.236 \times 10^{21}$  & 1.690  & $6.929 \times 10^{20}$  & 2.473  & $1.248 \times 10^{20}$  \\
 $\rho X(\isotm{He}{4})$     & $2.877 \times 10^{18}$  & 2.023  & $7.078 \times 10^{17}$  & 2.536  & $1.220 \times 10^{17}$  \\
 $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.947  & $4.402 \times 10^{16}$  & 2.222  & $9.436 \times 10^{15}$  \\
 $\rho X(\isotm{O}{16})$     & $1.688 \times 10^{14}$  & 1.658  & $5.349 \times 10^{13}$  & 1.950  & $1.385 \times 10^{13}$  \\
 $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.047  & $6.760 \times 10^{7}$   & 2.563  & $1.144 \times 10^{7}$   \\
\enddata
\end{deluxetable}

>>>>

\subsection{NSE Bailout}

There is one additional part of the time integration algorithm.
Because $T$ evolves during the reactions (for both Strang and SDC), it
is possible for a zone to start out not in NSE but evolve into NSE
during the reaction update.  When this happens, the ODE integrator may
fail, because an excessive number of timesteps is required.  Rather
than trigger our retry mechanism and throw away the entire timestep
and start over with a smaller $\Delta t$, we first check if the zone
evolved into NSE, and if so, we redo the zone update using the NSE
prescription.

@@

% run on 04-06-21
% Castro       git describe: 21.04-dirty
% AMReX        git describe: 21.03-58-g87c81d0ca
% Microphysics git describe: 21.04-1-g33a4d0fe

@@

% run on 04-06-21
% Castro       git describe: 21.04-dirty
% AMReX        git describe: 21.03-58-g87c81d0ca
% Microphysics git describe: 21.04-1-g33a4d0fe

\begin{deluxetable}{lllllll}
\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
\startdata
 $\rho$                      & $2.794 \times 10^{18}$  & 2.047  & $6.760 \times 10^{17}$  & 2.563  & $1.144 \times 10^{17}$  \\
 $\rho u$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
 $\rho v$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
 $\rho E$                    & $2.450 \times 10^{35}$  & 2.357  & $4.783 \times 10^{34}$  & 2.743  & $7.144 \times 10^{33}$  \\
 $\rho e$                    & $2.261 \times 10^{35}$  & 2.325  & $4.513 \times 10^{34}$  & 2.821  & $6.385 \times 10^{33}$  \\
 $T$                         & $2.236 \times 10^{21}$  & 1.690  & $6.929 \times 10^{20}$  & 2.473  & $1.248 \times 10^{20}$  \\
 $\rho X(\isotm{He}{4})$     & $2.877 \times 10^{18}$  & 2.023  & $7.078 \times 10^{17}$  & 2.536  & $1.220 \times 10^{17}$  \\
 $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.947  & $4.402 \times 10^{16}$  & 2.222  & $9.436 \times 10^{15}$  \\
 $\rho X(\isotm{O}{16})$     & $1.688 \times 10^{14}$  & 1.658  & $5.349 \times 10^{13}$  & 1.950  & $1.385 \times 10^{13}$  \\
 $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.047  & $6.760 \times 10^{7}$   & 2.563  & $1.144 \times 10^{7}$   \\
\enddata
\end{deluxetable}
